---
title: "H3K9me3 CHIPSeq rerun"
author: "Ola Hassan"
date: "1/4/20234"
output:
  pdf_document: default
  html_document: default
---

Read the bed files
```{r }

require(ChIPseeker)
require(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene
require(clusterProfiler)
library(org.Hs.eg.db)



peak_CT.1 <- readPeakFile("1_0GL8_028PBrown_siCT-1_H3K9me3_hg38_i81_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed")
covplot(peak_CT.1)

peak_CT.2 <- readPeakFile("2_0GL9_028PBrown_siCT-2_H3K9me3_hg38_i82_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed")
covplot(peak_CT.2)

peak_H7.1 <- readPeakFile("3_0GLA_028PBrown_siH7-1_H3K9me3_hg38_i85_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed")
covplot(peak_H7.1)

peak_H7.2 <- readPeakFile("6_0GVZ_028PBrown_siH7-2_H3K9me3_hg38_i70_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed")
covplot(peak_H7.2)


covplot(peak_CT.1, weightCol="V4")
covplot(peak_CT.2, weightCol="V4")
covplot(peak_H7.1, weightCol="V4")
covplot(peak_H7.2, weightCol="V4")


```


Heatmap of ChIP binding to TSS regions

```{r }

promoter <- getPromoters(TxDb=txdb, upstream=3000, downstream=3000)


tagMatrix_CT.1 <- getTagMatrix(peak_CT.1, windows=promoter)
tagMatrix_CT.2 <- getTagMatrix(peak_CT.2, windows=promoter)
tagMatrix_H7.1 <- getTagMatrix(peak_H7.1, windows=promoter)
tagMatrix_H7.2 <- getTagMatrix(peak_H7.2, windows=promoter)

tagHeatmap(tagMatrix_CT.1, xlim=c(-3000, 3000), color="red")
tagHeatmap(tagMatrix_CT.2, xlim=c(-3000, 3000), color="red")
tagHeatmap(tagMatrix_H7.1, xlim=c(-3000, 3000), color="red")
tagHeatmap(tagMatrix_H7.2, xlim=c(-3000, 3000), color="red")


#Average Profile of ChIP peaks binding to TSS region

plotAvgProf(tagMatrix_CT.1, xlim=c(-3000, 3000), 
            xlab="Genomic Region (5'->3') CT.1", ylab = "Read Count Frequency")

plotAvgProf(tagMatrix_CT.2, xlim=c(-3000, 3000), 
            xlab="Genomic Region (5'->3') CT.2", ylab = "Read Count Frequency")

plotAvgProf(tagMatrix_H7.1, xlim=c(-3000, 3000), 
            xlab="Genomic Region (5'->3') H7.1", ylab = "Read Count Frequency")

plotAvgProf(tagMatrix_H7.2, xlim=c(-3000, 3000), 
            xlab="Genomic Region (5'->3') H7.2", ylab = "Read Count Frequency")




```


Peak Annotation
```{r }
peakAnno_CT1 <- annotatePeak(peak_CT.1, tssRegion=c(-3000, 3000), 
                         TxDb=txdb, annoDb="org.Hs.eg.db")

peakAnno_CT2 <- annotatePeak(peak_CT.2, tssRegion=c(-3000, 3000), 
                         TxDb=txdb, annoDb="org.Hs.eg.db")

peakAnno_H71 <- annotatePeak(peak_H7.1, tssRegion=c(-3000, 3000), 
                         TxDb=txdb, annoDb="org.Hs.eg.db")

peakAnno_H72 <- annotatePeak(peak_H7.2, tssRegion=c(-3000, 3000), 
                         TxDb=txdb, annoDb="org.Hs.eg.db")

```



Visualize Genomic Annotation
```{r }
#adjust the margins for titles

par(mar=c(5,4,5,2)+0.2)
par(oma=c(0,0,2,0))
par(ps = 10, cex = 4, cex.main = 2) 
title(line=3, cex=1, 'main title')
par(mfrow=c(1,4), mar = c(5, 4, 4, 2) + 0.1)

plotAnnoPie(peakAnno_CT1, main='CT_1')
plotAnnoPie(peakAnno_CT2, main='CT_2')
plotAnnoPie(peakAnno_H71, main = 'H7_1')
plotAnnoPie(peakAnno_H72, main = 'H7_2')

par(mfrow = c(1,4))
plotAnnoBar(peakAnno_CT1)
plotAnnoBar(peakAnno_CT1)
plotAnnoBar(peakAnno_H71)
plotAnnoBar(peakAnno_H72)

#Since some annotation overlap, full annotation can show the overlap,
vennpie(peakAnno_CT1)
vennpie(peakAnno_CT2)
vennpie(peakAnno_H71)
vennpie(peakAnno_H72)

#Full annonatioation with the overlap
upsetplot(peakAnno_CT1)
upsetplot(peakAnno_CT2)
upsetplot(peakAnno_H71)
upsetplot(peakAnno_H72)

library(ReactomePA)

pathway_CT.1 <- enrichPathway(as.data.frame(peakAnno_CT1)$geneId)
head(pathway_CT.1)

pathway_CT.2 <- enrichPathway(as.data.frame(peakAnno_CT2)$geneId)
head(pathway_CT.2)

pathway_H7.1 <- enrichPathway(as.data.frame(peakAnno_H71)$geneId)
head(pathway_H7.1)

pathway_H7.2 <- enrichPathway(as.data.frame(peakAnno_H72)$geneId)
head(pathway_H7.1)

```


Compiling the 4 bedfiles in one list
```{r }
peaks <- as.list(c ("1_0GL8_028PBrown_siCT-1_H3K9me3_hg38_i81_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed",
             "2_0GL9_028PBrown_siCT-2_H3K9me3_hg38_i82_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed",
             "3_0GLA_028PBrown_siH7-1_H3K9me3_hg38_i85_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed",
           "6_0GVZ_028PBrown_siH7-2_H3K9me3_hg38_i70_uniqnorm_hits-W200-G600-FDR1E-10-island_filtered.bed" ))

peak=GenomicRanges::GRangesList(CT.1=readPeakFile(peaks[[1]]),
                                CT.2=readPeakFile(peaks[[2]]),
                                H7.1=readPeakFile(peaks[[3]]),
                                H7.2=readPeakFile(peaks[[4]]))



# the 4 peaks together
tagMatrixList <- lapply(peak, getTagMatrix, windows=promoter)
tagHeatmap(tagMatrixList, xlim=c(-3000, 3000),  color= "red")




plotAvgProf(tagMatrixList, xlim=c(-3000, 3000), 
            xlab="Genomic Region (5'->3') all peaks", ylab = "Read Count Frequency")

plotAvgProf(tagMatrixList, xlim=c(-3000, 3000), conf=0.95,resample=100, facet="row")

#peak annontation
peakAnnoList=lapply(peak, annotatePeak, TxDb=txdb, tssRegion=c(-3000, 3000), verbose=FALSE)

```

Visualize distribution of TF-binding loci relative to TSS for the 4 CHIPseq samples

```{r }
                      
plotAnnoBar(peakAnnoList)

plotDistToTSS(peakAnnoList)
```


Overlap between peak sets
```{r }
genes= lapply(peakAnnoList, function(i) as.data.frame(i)$geneId)
vennplot(genes)


```

Extract near by genes
```{r }

library(CompGO)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)

CT.1genes = annotateBedFromDb(gRanges=peak_CT.1, db=TxDb.Hsapiens.UCSC.hg38.knownGene)
CT.1genes = unlist(CT.1genes$gene_id)
head(CT.1genes)


CT.2genes = annotateBedFromDb(gRanges=peak_CT.2, db=TxDb.Hsapiens.UCSC.hg38.knownGene)
CT.2genes = unlist(CT.2genes$gene_id)
head(CT.2genes)

H7.1genes = annotateBedFromDb(gRanges=peak_H7.1, db=TxDb.Hsapiens.UCSC.hg38.knownGene)
H7.1genes = unlist(H7.1genes$gene_id)
head(H7.1genes)

H7.2genes = annotateBedFromDb(gRanges=peak_H7.2, db=TxDb.Hsapiens.UCSC.hg38.knownGene)
H7.2genes = unlist(H7.2genes$gene_id)
head(H7.2genes)


```

##Making an object from the genome wide annotation for Human.

```{r }
library(org.Hs.eg.db)
HA = org.Hs.eg.db

```

##Now annotate gene IDs to gene symbols and list those genes

```{r }

#CT1
CT.1genes = select(HA, keys = CT.1genes, columns = c("ENTREZID", "SYMBOL"), keytype = "ENTREZID")
head(CT.1genes)

CT.1genelist = unique(CT.1genes$SYMBOL)
head(CT.1genelist) #6925

#CT2
CT.2genes = select(HA, keys = CT.2genes, columns = c("ENTREZID", "SYMBOL"), keytype = "ENTREZID")
head(CT.2genes)

CT.2genelist = unique(CT.2genes$SYMBOL)
head(CT.2genelist) #7258



#H7.1
H7.1genes = select(HA, keys = H7.1genes, columns = c("ENTREZID", "SYMBOL"), keytype = "ENTREZID")
head(H7.1genes)

H7.1genelist = unique(H7.1genes$SYMBOL)
head(H7.1genelist) #6845


#H7.2
H7.2genes = select(HA, keys = H7.2genes, columns = c("ENTREZID", "SYMBOL"), keytype = "ENTREZID")
head(H7.2genes)

H7.2genelist = unique(H7.2genes$SYMBOL)
head(H7.2genelist) #7517

```


## Combine lists and remove duplications
```{r }
both_CT = intersect(CT.1genelist, CT.2genelist)
head(both_CT)


both_CT <- data.frame(both_CT)
head(both_CT) #6023 genes


both_H7 = intersect(H7.1genelist, H7.2genelist)
head(both_H7)

both_H7 <- data.frame(both_H7)
head(both_H7) #6138 genes

```


##Unique genes in both groups
```{r }

Unique_CT = setdiff(both_CT, both_H7)
head(Unique_CT) # 722 genes


Unique_H7 = setdiff(both_H7, both_CT)
head(Unique_H7) # 837 genes

write.csv(Unique_CT, "Gene mapping Unique for siCT and not siH7.csv", na="", row.names=F)
write.csv(Unique_H7, "Gene mapping Unique for siH7.csv", na="", row.names=F)


```






genes = lapply(peakAnnoList, function(i) as.data.frame(i)$geneId)
names(genes) = sub("_", "\n", names(genes))
compMF <- compareCluster(geneCluster   = genes, 
                         fun           = "enrichGO",
                         ont           = "MF",
                         OrgDb         ="org.Hs.eg.db",
                         pvalueCutoff  = 0.05, 
                         pAdjustMethod = "BH"
                         )
plot(compMF, showCategory = 20, title = "Molecular Function Enrichment")

genes = lapply(peakAnnoList, function(i) as.data.frame(i)$geneId)
names(genes) = sub("_", "\n", names(genes))
compKEGG <- compareCluster(geneCluster   = genes,
                         fun           = "enrichKEGG",
                         pvalueCutoff  = 0.05,
                         pAdjustMethod = "BH")
dotplot(compKEGG, showCategory = 15, title = "KEGG Pathway Enrichment Analysis")




```
